A Gravitationally Lensed Supernova with an Observable Two-Decade Time Delay

### 2020-06-23 Paper/Science/main.tex
Updates from Overleaf

We discovered \SNABC using data from the Hubble Space Telescope (HST) program {\it REsolved QUIEscent Magnified Galaxies} (REQUIEM, HST-GO-15663, PI:Akhshik). This project targets rare examples of quiescent galaxies that have been magnified by strong gravitational lensing. %These are typically too faint and compact to be resolved even with Hubble, but due to lensing are magnified. This provides a unique opportunity to  understand the still mysterious quenching mechnism. 
The brightest and most spectacular galaxy targeted by REQUIEM
is MRG-M0138, first discovered as a massive red galaxy (MRG) at $z=1.95$ \cite{newman_resolving_2018} behind the massive galaxy cluster MACS J0138.0-2155 \cite{ebeling_macs_2001}.
MRG-M0138 is quadruply lensed by the foreground cluster, which is at $z=0.338$.  
During analysis of REQUIEM observations obtained 13-14 July 2019 \cite{materials_methods_2020} we discovered three point 
sources that were present in archival HST images from 18-19 July 2016, part of the program 
that first confirmed the MRG-M0138 galaxy as a strongly-lensed object (HST-GO-14496; PI:Newman). 
Each point source is within 5 arcseconds of one of the four MRG-M0138 images.  None of the
three point sources are present in the REQUIEM HST data in 2019 (Fig \ref{fig:layout}). We infer that 
the point sources are multiple images of a single astrophysical 
transient in MRG-M0138, most likely a SN \cite{materials_methods_2020}.  
%Hereafter we call this multiply-imaged transient \SNABC.
The official designation for this multiply-imaged astrophysical transient is \SNABC.

@@

%BibTeX users: After compilation, comment out the following two lines and paste in
% the generated .bbl file. 
\clearpage
\begin{figure*}
    \centering
    \includegraphics[draft=False,width=0.9\textwidth]{Paper/Figures/fig1_layout_sn4.pdf}
    \caption{Overview of the MACSJ0138 cluster field and the locations of the \SNABC images  (SN1--3) and its multiply-imaged host galaxy (H1--4). The wide-field view in $a)$ is 40\arcsec\ on a side with ticks indicating 10\arcsec\ intervals.  Panels \emph{b}--\emph{i} show 4\arcsec\ cutouts around the lensed SN images with 1\arcsec\ ticks.  The three-color images are generated from the WFC3/IR filters as indicated; panels \emph{bcde} show the imaging from 2016 July where the SN was visible and \emph{fghi} show the later imaging from 2019 July where the the SN has faded away.  Panels \emph{ei} indicate the location of the fourth image predicted to appear in $\approx$2037. All panels use the late-epoch F125W imaging for the green channel; nevertheless, it is immediately clear that the SN2 image is substantially bluer than the other two, helps to constrain the relative age of each SN image and the transient classification as a likely Type Ia supernova explosion.}

@@

\begin{table}[]
    \centering
    \begin{tabular}{rccl}
     ID &   R.A. (deg) & Dec. (deg) & $z$ \\
     \midrule
H1 & 24.5099018 & $-$21.9260130 & 1.95 \\
H2 & 24.5132090 & $-$21.9299032 & 1.95 \\
H3 & 24.5164138 & $-$21.9303172 & 1.95 \\
H4 & 24.5176117 & $-$21.9233433 & 1.95 \\
SN1 & 24.5100753 & $-$21.9273418 & 1.95 \\
SN2 & 24.5123198 & $-$21.9297875 & 1.95 \\
SN3 & 24.5151253 & $-$21.9306659 & 1.95 \\
3.1 & 24.5169659 & $-$21.9234814 & 0.7663 \\
3.2 & 24.5161061 & $-$21.9231225 & 0.7663 \\
3.3 & 24.5184361 & $-$21.9264250 & 0.7663 \\
3.4 & 24.5146833 & $-$21.9261020 & 0.7663 \\
    \end{tabular}
    \caption{Multiple images used as constraints in our parametric model. From left to right: image ID, right ascension, declination, spectroscopic redshift.  Coordinates are in the J2000 reference frame.}
    \label{tab:mulimages}
\end{table}

>>>>

We discovered \SNABC using data from the Hubble Space Telescope (HST) program {\it REsolved QUIEscent Magnified Galaxies} (REQUIEM, HST-GO-15663, PI:Akhshik). This project targets rare examples of quiescent galaxies that have been magnified by strong gravitational lensing. %These are typically too faint and compact to be resolved even with Hubble, but due to lensing are magnified. This provides a unique opportunity to  understand the still mysterious quenching mechnism. 
The brightest and most spectacular galaxy targeted by REQUIEM
is MRG-M0138, first discovered as a massive red galaxy (MRG) at $z=1.95$ \cite{newman_resolving_2018} behind the massive galaxy cluster MACS J0138.0-2155 \cite{ebeling_macs_2001}.
MRG-M0138 is quadruply lensed by the foreground cluster, which is at $z=0.338$.  
During analysis of REQUIEM observations obtained 13--14 July 2019 \cite{materials_methods_2020} we discovered three point 
sources that were present in archival HST images from 18--19 July 2016, part of the program 
that first confirmed the MRG-M0138 galaxy as a strongly-lensed object (HST-GO-14496; PI:Newman). 
Each point source is within 5 arcseconds of one of the four MRG-M0138 images.  None of the
three point sources are present in the REQUIEM HST data in 2019 (Fig \ref{fig:layout}). We infer that 
the point sources are multiple images of a single astrophysical 
transient in MRG-M0138, most likely a SN \cite{materials_methods_2020}.  
%Hereafter we call this multiply-imaged transient \SNABC.
The official designation for this multiply-imaged astrophysical transient is \SNABC.

@@

%BibTeX users: After compilation, comment out the following two lines and paste in
% the generated .bbl file. 
\clearpage
\begin{figure*}
    \centering
    \includegraphics[draft=False,width=0.9\textwidth]{Paper/Figures/fig1_layout_sn4.pdf}
    \caption{Overview of the MACSJ0138 cluster field and the locations of the \SNABC images  (SN1--3) and its multiply-imaged host galaxy (H1--4). The wide-field view in $a)$ is 40\arcsec\ on a side with ticks indicating 10\arcsec\ intervals.  
    %Labels marked 3.1--3.4 indicate the locations of a separate  multiply-imaged [O\textsc{ii}]-emitter at $z=0.77$ used to help constrain the lensing potential. 
    Panels \emph{b}--\emph{i} show 4\arcsec\ cutouts around the lensed SN images with 1\arcsec\ ticks.  The three-color images are generated from the WFC3/IR filters as indicated; panels \emph{bcde} show the imaging from 2016 July where the SN was visible and \emph{fghi} show the later imaging from 2019 July where the the SN has faded away.  Panels \emph{ei} indicate the location of the fourth image predicted to appear in $\approx$2037. All panels use the late-epoch F125W imaging for the green channel; nevertheless, it is immediately clear that the SN2 image is substantially bluer than the other two, helps to constrain the relative age of each SN image and the transient classification as a likely Type Ia supernova explosion.}

@@

\begin{table}[]
    \centering
    \begin{tabular}{rccl}
     ID &   R.A. (deg) & Dec. (deg) & $z$ \\
     \midrule
H1 & 24.5099018 & $-$21.9260130 & 1.95 \\
H2 & 24.5132090 & $-$21.9299032 & 1.95 \\
H3 & 24.5164138 & $-$21.9303172 & 1.95 \\
H4 & 24.5176117 & $-$21.9233433 & 1.95 \\
SN1 & 24.5100753 & $-$21.9273418 & 1.95 \\
SN2 & 24.5123198 & $-$21.9297875 & 1.95 \\
SN3 & 24.5151253 & $-$21.9306659 & 1.95 \\
3.1 & 24.5169659 & $-$21.9234814 & 0.7663 \\
3.2 & 24.5151039 & $-$21.9231406 & 0.7663 \\
3.3 & 24.5184361 & $-$21.9264250 & 0.7663 \\
3.4 & 24.5146833 & $-$21.9261020 & 0.7663 \\
    \end{tabular}
    \caption{Multiple images used as constraints in our parametric model. From left to right: image ID, right ascension, declination, spectroscopic redshift.  Coordinates are in the J2000 reference frame.}
    \label{tab:mulimages}
\end{table}
